---
title: "Playing Time Optimization"
output: html_document
date: "2024-11-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DT)

source("~/Documents/Research/Analytics%20Dev/Working/KTENENBAUM/Roster_Optimization/Position_Players/00_Process_Data.R")
source("~/Documents/Research/Analytics%20Dev/Working/KTENENBAUM/Roster_Optimization/Position_Players/01_Simulate_Seasons.R")
source("~/Documents/Research/Analytics%20Dev/Working/KTENENBAUM/Roster_Optimization/Position_Players/Roster_Optimization.R")

source("~/Documents/Research/Analytics%20Dev/Working/KTENENBAUM/Roster_Optimization/Pitching/00_Load_Pitching_Data.R")
source("~/Documents/Research/Analytics%20Dev/Working/KTENENBAUM/Roster_Optimization/Pitching/01_Piching_Optimization.R")
source("~/Documents/Research/Analytics%20Dev/Working/KTENENBAUM/Roster_Optimization/Pitching/02_Simulate_Pitching.R")

### Position Players
players <- c(
    "920aebff-8966-4696-aa37-275a785fd06b" # Taylor Ward
  , "38147acb-e01f-4dba-bd59-5fa9adf7834e" # Alec Bohm
  , "eb7c0429-883e-4e38-b76e-3bd894e136f4" # Yandy Diaz
  , "72c66910-6dd5-4810-95cc-f244faea5d2c" # Brandon Lowe
  , "d2b8d21c-3359-43eb-a39b-89d21a4bfa63" # Nate Lowe
  , "72f94915-e477-4e2e-b2c8-b7febe2484a2" # Brandon Marsh
  , "1b6e4920-d730-42b3-8f0d-4f1559638cfa" # Ryan Mountcastle
  , "0c329903-4730-4c20-bd5c-de6d7b599ee7" # Josh Naylor
  , "da6c5b56-34b2-4bd0-a417-6c9da345e58a" # Luis Rengifo
  , "fdaf8b60-1165-42b6-b3b1-d2340ff42dbb" # Pavin Smith
  , "1c40ff76-048b-4ff2-914a-70a0856f0f06" # Bryson Stott
  , "82fbd839-c165-45bc-81fb-51bc52a1d12f" # Lane Thomas
  , "a31d7683-ce8d-429b-9fbe-dab5d1dcfe34" # Lamonte Wade
)


addPlayerPositions <- data.frame(
  PiratesID = c("2dac7862-21dd-4926-8452-075757d9cccb", "50ec0fe4-8960-4188-a1a3-36aae400e1e3", "48b0cda6-abea-40ec-8120-c7ee12ceb3e6"),
  PositionTypeID = c(3, 8, 8),
  FRAA = c(-2, -7, 5)
)

playerConstraints <- data.frame(
  PiratesID = "48b0cda6-abea-40ec-8120-c7ee12ceb3e6",
  Position = '6',
  Operator = "<=",
  RHS = 0
)

adjPlayerOffense <- data.frame(
  PiratesID = c("72f94915-e477-4e2e-b2c8-b7febe2484a2","72f94915-e477-4e2e-b2c8-b7febe2484a2"),
  split = c('vL', 'vR'),
  adj = c(7, -3)
)




base <- optimize_rosters_agg_all(df_repeated, Org = 'PIT'
                                 , playerRemovals = c('fd3dc6b8-92c4-46e2-9417-7a55e7c0ff83')
                                 # , addPlayerPositions = data.frame(
                                 #   PiratesID = c("2dac7862-21dd-4926-8452-075757d9cccb", "50ec0fe4-8960-4188-a1a3-36aae400e1e3"),
                                 #   PositionTypeID = c(3, 8),
                                 #   FRAA = c(-2, -7)
                                 # )
                                 , addPlayerPositions = addPlayerPositions
                                 , playerConstraints = playerConstraints
                                 # , adjPlayerOffense = adjPlayerOffense
                                 )


# write.csv(allBase, "basePT.csv", row.names = FALSE)



  
  roster <- optimize_rosters_agg_all(df_repeated, Org = 'PIT'
                           , playerAdditions = c(#'fd3dc6b8-92c4-46e2-9417-7a55e7c0ff83',
                                                 "920aebff-8966-4696-aa37-275a785fd06b")
                           , addPlayerPositions = addPlayerPositions
                           , playerConstraints = playerConstraints)
  
  
  marginalGain <- roster$obj - base$obj
  
  pt_diff <- roster$playing_time %>% filter(split == 'vB') %>% inner_join(base$playing_time %>% filter(split == 'vB'), by = c('PlayerName','PiratesID', 'split')) %>% 
    mutate(Diff = Total.x - Total.y)  %>% arrange((Diff)) %>% select(PiratesID, PlayerName, Playing_Time_new = Total.x, Playing_Time_old = Total.y, Diff) %>% 
    mutate(diff_text = paste0(PlayerName, round(Diff*100), "%"))
  
  # movers <- paste0(pt_diff %>% filter(abs(Diff) > .05) %>% top_n(5) %>% pull(diff_text), collapse = ', ')
  


# output_frame <- data.frame(
#     PlayerName = unlist(map(player_output, "PlayerName"))
#   , MarginalRuns = unlist(map(player_output, "runs"))
#   , BiggestMovers = unlist(map(player_output, "movers"))
#   
# )

PP_Wins <- read.csv('Org_Win_Probs.csv')
Orgs_Pitching <- read.csv("Playoff_Odds/PitchingWins.csv")
playoff_cond_probs <- read.csv('Playoff_Odds/playoff_probs.csv')
playoffFit <- readRDS("playoffFit_PIT_2025.RDS")

win_projections <- PP_Wins %>% inner_join(Orgs_Pitching, by = 'Org') %>% 
  mutate(Win_Total = Wins + WAA.y) %>% 
  select(
    Org, Wins, WAA.y, Win_Total
  )

base_PIT_win <- win_projections %>% filter(Org == 'PIT') %>% pull(Win_Total)

findWinDist <- function(Wins, playoff_cond_probs, playoffFit, sampleVar = 6.25, talentVar = 3){
  totalVar <- sqrt(sampleVar^2 + talentVar^2)
  winVals <-65:110
  winFrame <- data.frame(
    W = winVals,
    prob = pnorm(winVals +.5, mean = Wins, sd = totalVar) - pnorm(winVals -.5, mean = Wins, sd = totalVar)
  ) 

  winFrame$Playoff = predict(playoffFit, newdata = winFrame, type = 'response')
    
  
  winFrame <- bind_cols(winFrame, apply(winFrame, 1, function(x) x['Playoff']*playoff_cond_probs) %>% bind_rows()) 
  
  winFrame %>% summarise(
    Wins = Wins,
    Playoff = sum(prob*Playoff)/sum(prob),
    WCWin   = sum(prob*WCWin)/sum(prob),
    DSWin = sum(prob*DSWin)/sum(prob),
    LCSWin = sum(prob*LCSWin)/sum(prob),
    WSWin = sum(prob*WSWin)/sum(prob),
  ) %>% return()
}


### Pitchers
pitching_base <- optimize_pitching_agg(df_repeated_pitching, Org = "PIT")

# addPitcher <- "44c8dce1-bb49-4d23-91c8-a4194afc7b19" # Camilo Doval
addPitcher <- "fcbbe69a-e6ea-464d-ade7-76606878be83" # Reid Detmers
pitching_add <- optimize_pitching_agg(df_repeated_pitching, Org = "PIT", playerAdditions = addPitcher)

```

You can see our base playing time distribution here with combined handedness followed by the same thing vL and vR. Note that we have added the following constraints and adjustments to our roster. These are all adjustable and we are doing these more as a demonstration that we can. 
* Brian Reynolds can play 1B as a -2 defender.
* Billy Cook is a -7 CF
* Oneil Cruz is a +5 CF
* Andrew McCutchen is on the roster

```{r, echo=FALSE}
brks <- seq(from = 0, to = 1, by = .1)
clrs <- round(seq(255, 40, length.out = length(brks) + 1), 0) %>%
  {paste0("rgb(255,", ., ",", ., ")")}

datatable(base$playing_time %>% ungroup() %>% filter(split == 'vB') %>% 
            select(-PiratesID) %>% relocate(split, .after = PlayerName) %>% 
            arrange(desc(Total))) %>%
  formatPercentage(3:12) %>% 
  formatStyle(names(base$playing_time)[3:12], backgroundColor = styleInterval(brks, clrs))


```
```{r, echo=FALSE}
brks <- seq(from = 0, to = 1, by = .1)
clrs <- round(seq(255, 40, length.out = length(brks) + 1), 0) %>%
  {paste0("rgb(255,", ., ",", ., ")")}

datatable(base$playing_time %>% ungroup() %>% filter(split == 'vL') %>% 
            select(-PiratesID) %>% relocate(split, .after = PlayerName) %>% 
            arrange(desc(Total))) %>%
  formatPercentage(3:12) %>% 
  formatStyle(names(base$playing_time)[3:12], backgroundColor = styleInterval(brks, clrs))


```
```{r, echo = FALSE}
brks <- seq(from = 0, to = 1, by = .1)
clrs <- round(seq(255, 40, length.out = length(brks) + 1), 0) %>%
  {paste0("rgb(255,", ., ",", ., ")")}

datatable(base$playing_time %>% ungroup() %>% filter(split == 'vR') %>% 
            select(-PiratesID) %>% relocate(split, .after = PlayerName) %>% 
            arrange(desc(Total))) %>%
  formatPercentage(3:12) %>% 
  formatStyle(names(base$playing_time)[3:12], backgroundColor = styleInterval(brks, clrs))


```


We would expect this team to have the following outcomes.

```{r, echo=FALSE}

winDist1 <- findWinDist(Wins = win_projections %>% filter(Org == 'PIT') %>% pull(Win_Total), 
            playoff_cond_probs = playoff_cond_probs,
            playoffFit = playoffFit) %>% mutate(Wins = round(Wins,1))

winDist1 %>% mutate(Type = "Baseline") %>% relocate(Type) %>% datatable() %>%
  formatPercentage(2:ncol(winDist1)+1) 

```

Now, we can add in other players and see how playing time and outcomes might change. Let's look at a world where we add Yandy Diaz to our roster. The playing time changes are below:

```{r, echo=FALSE}
brks <- seq(from = 0, to = 1, by = .1)
clrs <- round(seq(255, 40, length.out = length(brks) + 1), 0) %>%
  {paste0("rgb(255,", ., ",", ., ")")}

datatable(roster$playing_time %>% ungroup() %>% filter(split == 'vB') %>% 
            select(-PiratesID) %>% relocate(split, .after = PlayerName) %>% 
            arrange(desc(Total))) %>%
  formatPercentage(3:12) %>% 
  formatStyle(names(base$playing_time)[3:12], backgroundColor = styleInterval(brks, clrs))


```

```{r, echo = FALSE}
brks <- seq(from = 0, to = 1, by = .1)
clrs <- round(seq(255, 40, length.out = length(brks) + 1), 0) %>%
  {paste0("rgb(255,", ., ",", ., ")")}

datatable(roster$playing_time %>% ungroup() %>% filter(split == 'vL') %>% 
            select(-PiratesID) %>% relocate(split, .after = PlayerName) %>% 
            arrange(desc(Total))) %>%
  formatPercentage(3:12) %>% 
  formatStyle(names(base$playing_time)[3:12], backgroundColor = styleInterval(brks, clrs))


```

```{r, echo=FALSE}
brks <- seq(from = 0, to = 1, by = .1)
clrs <- round(seq(255, 40, length.out = length(brks) + 1), 0) %>%
  {paste0("rgb(255,", ., ",", ., ")")}

datatable(roster$playing_time %>% ungroup() %>% filter(split == 'vR') %>% 
            select(-PiratesID) %>% relocate(split, .after = PlayerName) %>% 
            arrange(desc(Total))) %>%
  formatPercentage(3:12) %>% 
  formatStyle(names(base$playing_time)[3:12], backgroundColor = styleInterval(brks, clrs))


```

This roster would be expected to have the following outcomes:
```{r, echo=FALSE}


utilityFunction <- c(Win = 1, Playoff = 15, winWC = 25, winDS = 40, winLCS = 60, winWS = 85)


winDist <- findWinDist(Wins = win_projections %>% filter(Org == 'PIT') %>% pull(Win_Total) + marginalGain/10, 
            playoff_cond_probs = playoff_cond_probs,
            playoffFit = playoffFit) %>% mutate(Wins = round(Wins,1))

win_comparison <- bind_rows(
  winDist1 %>% mutate(Type = "Baseline") %>% relocate(Type),
  winDist %>% mutate(Type = 'w/ Yandy') %>% relocate(Type)
  )



win_comparison %>% datatable() %>%
  formatPercentage(3:(ncol(winDist)+1), digits = 1) 
# winDist %>% datatable() %>%
  # formatPercentage(2:ncol(winDist)) 

```

We can do the same thing with pitching and compare our pitching allocation and team outcome projections for our team with and without Camilo Doval. Note, that I am not yet including injuries and I am making some very irresponsible assumptions about pitching. These should look far from perfect and are more to give you the idea for how this can work rather than be a fully functioning tool.

### Base team pitching allocation

```{r, echo = FALSE}

pitching_base$playing_time %>% ungroup() %>% select(PlayerName, SP, IP, LI, Lev_Innings) %>% 
  mutate(IP = round(IP, 1),
         LI = round(LI, 2),
         Lev_Innings = round(Lev_Innings, 1)) %>%
  arrange(desc(Lev_Innings)) %>% 
  datatable() %>% 
  formatPercentage(c(2), digits = 1) 


```
We can compare that pitching allocation to the one with Doval added
### Camilo Doval Added
```{r, echo = FALSE}

pitching_add$playing_time %>% ungroup() %>% select(PlayerName, SP, IP, LI, Lev_Innings) %>% 
  mutate(IP = round(IP, 1),
         LI = round(LI, 2),
         Lev_Innings = round(Lev_Innings, 1)) %>%
  arrange(desc(Lev_Innings)) %>% 
  datatable() %>% 
  formatPercentage(c(2), digits = 1) 


```

We can also see the impact of Doval on our team's outcome probabilities:

```{r, echo = FALSE}

marginalGain_pitching <- -(pitching_add$obj - pitching_base$obj)/10/9
winDist2 <- findWinDist(Wins = win_projections %>% filter(Org == 'PIT') %>% pull(Win_Total) + marginalGain_pitching, 
            playoff_cond_probs = playoff_cond_probs,
            playoffFit = playoffFit) %>% mutate(Wins = round(Wins,1))

winDist2 <- winDist2 %>% mutate(Type = 'w/ Doval') %>% relocate(Type)

win_comparison <- bind_rows(win_comparison, winDist2)

win_comparison %>% datatable() %>%
  formatPercentage(3:(ncol(winDist)+1), digits = 1) 

```

We can actually take this a step further and try to define our team's utility. Let's assume that we want to value each of these outcomes as below:


 Outcome | Utility 
|-------|------|
| Marginal Win | `r utilityFunction['Win']` |
| Playoffs | `r utilityFunction['Playoff']` |
| win Wild Card Series | `r utilityFunction['winWC']` |
| win Div Series | `r utilityFunction['winDS']` |
| win LCS | `r utilityFunction['winLCS']` |
| win World Series | `r utilityFunction['winWS']` |

Using these values, we can look at how much team utility we gain by adding Yandy Diaz or Camilo Doval to our roster. While we still need to figure out how to value utility in dollars, I think we can find ways to do that based on the price of buying it on the market at any given time. 
```{r, echo = FALSE}
win_comparison$Utility <- apply(win_comparison[,-1], 1, function(x) x * utilityFunction) %>% colSums() %>% round()

win_comparison %>% datatable() %>%
  formatPercentage(3:(ncol(winDist)+1), digits = 1) 

```

I just picked out these values of utility, so I wanted to explore how it's working a bit more. Here is how our team's utiilty changes based on our projected win total. You can also see how they might change if we value advancement in the playoffs more or less. 
```{r, echo = FALSE}
winVec <- 65:100

# Baseline
utility_curve_base <- sapply(winVec, function(w){
  winDist <- findWinDist(Wins = w, 
            playoff_cond_probs = playoff_cond_probs,
            playoffFit = playoffFit) %>% mutate(Wins = round(Wins,1))
  
  (winDist* utilityFunction) %>% as.matrix() %>% t() %>%  sum() %>% round() %>% return()
  
})

utilityFrame <- data.frame(
  Wins = winVec,
  Utility = utility_curve_base
)


# Value Playoffs

utilityFunction_high <- utilityFunction
utilityFunction_high[2:length(utilityFunction)] <- utilityFunction_high[2:length(utilityFunction)]*5.5
  

utility_curve_playoff_high <- sapply(winVec, function(w){
  winDist <- findWinDist(Wins = w, 
            playoff_cond_probs = playoff_cond_probs,
            playoffFit = playoffFit) %>% mutate(Wins = round(Wins,1))
  
  (winDist* utilityFunction_high) %>% as.matrix() %>% t() %>%  sum() %>% round() %>% return()
  
})

utilityFrame <- utilityFrame %>% 
  mutate(
    Utility_Playoff_High = utility_curve_playoff_high
  )

utilityFrame <- utilityFrame %>% 
  mutate(Utility = scale(Utility),
         Utility_Playoff_High = scale(Utility_Playoff_High))


plotFrame <- bind_rows(
  utilityFrame %>% select(Wins, Utility = Utility) %>% mutate(type = 'Baseline'),
  utilityFrame %>% select(Wins, Utility = Utility_Playoff_High) %>% mutate(type = 'Reward Playoffs')
)


ggplot(data = plotFrame) + geom_line(aes(x = Wins, y = Utility, colour = type)) 

```
